paper1_figures

### 2023-11-06 scoping_review_short.Rmd
short version of all the factors

-   **try to shorten as much as possible the result section by putting detailed results on the table**

-   **put the references in the table (ex: gender, 11 papers --\> put the 11 references in the case)**

@@

Five individual factors that can impact cooperation emerged from the litterature review: gender, social status, group identification, values and personality traits, and trust. In the following sections, the principal effects of these factors on cooperation are described. Moreover, an overview of the moderators of each effect is provided. More details about moderators are included in the Annexe A.

@@

Social identity has a strong, positive effect on cooperation: the more individuals identify with a group, the more they cooperate with other members of the group [@VanVugt2009], for instance by providing more voluntary activities for the collective good [@Noonan2016]. In fact, humans have a very developed sense of belonging to social groups, and when they strongly identify with a social group, they are more concerned about their group's well-being and reputation [@VanVugt2009] and they accord more importance to group-related goal [@Arora2016]. A strong identification also maintains high levels of cooperation over time [@Zhang2019]. Finally, further confirmations of the strength of this effect are the fact that it is the same for individualist and collectivist people [@Chen2007] and that group identification can be easily implemented in a laboratory experiment and still have a strong effect on cooperation [@Wit1992].

@@

Intragroup cooperation can be enhanced via a competition between different groups, because in such a context individual's and group's interests are aligned [@Puurtinen2009]. Moreover, a study showed that intergroup competition increased cooperation, while individual, dispositional competitiveness had the opposite effect [@Nockur2020]. This demonstrated that the increase of cooperation via competition is due to group membership factors, and not to competition in itself.

Despite the beneficial effects of inducing an intergroup competition, this could be harmful in real-life situations, like for instance in case of resources shared between several groups. In fact, there is a greater risk of over-exploitation if an intergroup competition is present [@VanVugt2009]. To maintain the positive effect of group identification while eliminating the risks associated to the competition, it is useful to prime a superordinate identity [@VanVugt2009]. This can be done by making salient the common characteristics of the different groups instead of their differences [@Kramer1984] or by inducing the feeling of a common fate [e.g. a collective risk, @Zhang2019].

@@

@@

Cooperation rates can be modified by introducing incentives, who can be either punishments or rewards[^1]. In the literature, punishment has received a greater attention than rewards [@Balliet2011a; @VanDijk2015], and its positive effect on cooperation is very consensual. In general, when there is the possibility for participants to punish others members based on their contributions, cooperation is higher than when this is not possible [e.g. @Pfattheicher2018; @Balliet2011a]. Moreover, the presence of a punishment system increases the perception of fairness [@Nockur2021].

@@

However, this main results is subjected to several moderators. First, the punishment has a bigger impact on cooperation if its implementation is costly for the punishers than if it is free [@Balliet2011a], but at the same time the cost should not be too high [@Egas2008]. Moreover, punishments are more effective if the interaction is among friends than among strangers, and if people know that they will have to interact repeatedly with each other [@Balliet2011a]. In fact, in short interactions, if punishment is costly, individuals who punish others are the ones with the lowest personal income [@Milinski2012], and this could discourage the application of the fine. Also, an experiment showed that a punishment system increased cooperation in groups where all members had the same amount of resources, but not in groups where income was heterogeneous [@Kingsley2016]. In fact, @Chen2022 found that, in groups where the initial endowment was heterogeneous, participants that received less didn't punish the others "low-status" members. This solidarity could therefore encourage free-riding. Moreover, the effect of punishment is stronger when only cooperators have the power to punish defectors, than when everybody can punish [@Xiao2016]. Finally, a plain situation is also needed to have a stronger effect of punishment: the link between behaviour and outcome must be clear, without random factors affecting the final outcome [@Xiao2016], and the punishment must be highly effective [@Egas2008].

However, punishment must be applied with carefulness: @Rockenbach2019 compared the strategies of different groups, and found that the most successful ones (whose members had the highest payoff at the end of the experiment) punished with caution and not in early stages of the game. In these groups, a norm of when and how punishment should be applied seems to exist. Moreover, punishment has also two main negative sides. First, it could entails a second-order social dilemma [@Milinski2012], since cooperative participants can free-ride on the effort of other cooperators who punish non-cooperators. Secondly, anti-social punishment can arise. This term refer to the punishment of cooperators by non-cooperators, and it is present when all participants can punish independently of their contributions [@Xiao2016]. In this context, defectors can see cooperators as rule breakers, and anti-social punishment is a sort of retaliation [@VanDijk2015].

A possible solution could be to implement a system of democratic punishment, where a fine is given only if a majority of participants agree. In fact, compared to a peer punishment system where participants decide individually if and who punish, in a democratic system anti-social punishment is less present, general punishment level is lower, and final payoff and cooperation rates are higher [@Pfattheicher2018].

Rewarding systems, on the other side, are less investigated. Globally, they have a positive effect on cooperation [@Balliet2011a]. However, some studies found that they were more efficient than punishment [e.g. @Chen2022], while others came to the opposite conclusion [e.g. @Milinski2012] or found no difference between the two systems [e.g. @Balliet2011a]. In a meta-analysis, @Balliet2011a found that rewards had a stronger effect on cooperation if their implementation was costly rather than free, and in iterated, rather than single, interactions. Moreover, rewarding was more effective in situations where participants had to decide how much take from a common resource, compared to when they had to decide how much contribute to a common good.

In general, incentives seem to be a powerful tool to enhance cooperation, but they have a dark side. First, they reduce participants' intrinsic motivation [@VanDijk2015], by framing the situation as an economic, rather than ethical, dilemma [@VanVugt2009]. In fact, @Schlosser2018 demonstrated that people's level of justice sensitivity could predict their contributions to a public good game, but this link disappeared when a punishment system was introduced. Secondly, when incentives are present people think that others cooperate just because of them, and this undermines their trust in other members, and especially the trust in the fact that others are internally motivated to cooperate. This reduced trust in others drops cooperation levels if incentives are removed [@Irwin2014; @Mulder2006].

Finally, incentives are often studied in economic formats, but they can also be social. For example, the expectation of honour or shame (after cooperating or defecting, respectively) is enough to enhance cooperation [@Jacquet2011]. Moreover, expressing disapproval toward defectors can boost cooperation by inducing more compliance to the cooperative social norm [@Nelissen2013]. Even an ostracism system where participants can vote to exclude other members from the game is very effective in increasing cooperation [@Maier-Rigaud2010]. In general, social sanctions can be even more efficient than economic ones, because cooperation do not drop after their removal [@Nelissen2013].

To conclude, economic incentives can have a positive effect on cooperation, but their efficient implementation must be insured all along the interaction, otherwise cooperation rates would drop. On the other hand, social incentives seem to have the same positive effect, and their removal do not cause a reduction of cooperation. However, there is a lack of studies focusing on this kind of sanctions.

@@

Communication has a global positive effect on cooperation [e.g. @Balliet2010; @Bicchieri2002], but this effect is subjected to several moderators. First, talking is more powerful than sending written messages, because it allows non-verbal communication and facilitates the social norms formation process. Moreover, communication's effect is stronger in large groups: when many people are present cooperation rates are lower, so communication between members can buffer this negative effect [@Balliet2010]. Finally, if communicating with others requires a personal cost, the effect is weaker than if communication is free [@Janssen2014].

The content of the messages is also essential. @Koessler2021 identified four steps necessary for an effective communication in social dilemmas: problem awareness, identification of strategies, agreement and ratification. In another study, groups who used communication to foster a common social identity, promote cooperative social norms and ensure that all the players publicly committed displayed higher rates of cooperation than other groups [@Adams2022]. However, casual chit-chat also plays a very important role in increasing cooperation [@Koessler2021]. Also, the increase of cooperation through communication can be explained by the normative content that people are shaping while exchanging messages [@Janssen2014], and by the fact that it allows participant to reach a consensus about the best strategy [@Hopthrow2005]. Moreover, even if communication also helps to create a social identity, the effect of social norms is a stronger mediator of the link between communication and cooperation [@Bicchieri2002].

Another factor that influences the effect of communication on cooperation is whether individuals exchange messages before, during or after the cooperative phase. Concerning the post-game communication, @Kumakawa2013 found that receiving a feedback about the cooperation choice made previously can increase cooperation in the following interactions, especially if the feedback is negative. However, just the fact of knowing that a discussion will follow is not enough to enhance cooperation, and can even have a counterproductive effect, because forcing people to think about a future discussion make them reflect more on the situation and thus act more rationally [@Torsvik2011]. On the other hand, communication before or during the interaction have both a positive impact on cooperation, and the difference between their effect is not significant [@Balliet2010].

Finally, several studies investigated the effect of gossip as a form of communication that can increase cooperation. Gossip is defined as a spread of reputational information about another person [@Feinberg2014], and it increases cooperation in two ways. First, if people know that gossip is possible, they tend to cooperate more to avoid the risk of being the target of such gossip and therefore of being excluded from following interactions. Secondly, having information about the past behaviour of potential partners can help to select the most cooperative individuals in an interaction situation [@Feinberg2014]. Moreover, the same authors found that pro-social individuals gossiped more than pro-self. This shows that gossip can have the pro-social function of protecting others from a potential exploitation [@Feinberg2012]. However, gossip has also some self-interest effects: for instance, observing another subject being exploited and having the opportunity to gossip about the exploiter significantly reduced negative affects in @Feinberg2012 's participants. Moreover, providing correct information while gossiping helps to build a positive reputation [@Giardini2021].

In conclusion, when people have the possibility to communicate before, during or after a cooperative interaction, the cooperation level increases. Even if the strength of this effect can vary depending on the situation, communication never seems to have a negative impact on cooperation.

@@

Social norms can provide a guide on how to behave, especially when there is uncertainty about others' behaviour [@VonBorgstede2018]. @Thogersen2008 studied the effect of social norms on daily social dilemma situations such as environmental behaviours, and found that both injunctive (what others approve) and descriptive (what others do) norms are positively related to behaviours. Moreover, cooperation can be maintained via social norms, because people who do not respect them are susceptible of social sanctions. In fact, if social approval is possible, people cooperate more [@Rege2004].

There are several explanations of social norms' effect on cooperation. First, according to the appropriateness framework, when taking a decision people ask themselves: "What does a person like me do in a situation like this?" [@Arora2012]. Therefore they need to define the situation as a cooperative task and to strongly identify with the group [@VanLange2013]. This leads to a greater accessibility of group's norms and consequently (if the norms are cooperative) to more cooperation. Testing this assumption, @Arora2012 showed that a greater identification with social group lead to higher levels of cooperation. The second explanation concerns conformity: in general people adapt their cooperation to others' behaviour, but this is usually explained by the reciprocity principle (cooperation enhances the probability that others will do the same, and therefore that the final payoff will be higher) [@Bardsley2005a]. However, these authors demonstrated that people conform to others' behaviour even if this has no impact on their personal payoff. Therefore, people have a general tendency to imitate others to follow social norms.

Finally, social norms can be modified by participants. For instance, @Weber2008 demonstrated that in several social dilemma experiment there are consistent cooperators, i.e. people who always cooperate. The presence of these people can increase other members' cooperation, and this effect is mediated by the perception of the social norms. Moreover, social norms can also be shaped via communication [@Bicchieri2002].

In conclusion, social norms have a significant effect on people's cooperation in social dilemma situations. Since they are quite easily modified [e.g. @Weber2008], shaping cooperative norms can be a powerful instrument to improve cooperation.

@@

There are not many studies on the influence of anonymity on cooperation, but, according to @Wang2017, most of the factors that can increase cooperation rely on a reduction of anonymity. For instance, it is not possible to punish defectors without knowing who they are. Therefore, in general it seems that non-anonymous situations can significantly increase cooperation and reduce defection [@Wang2017]. This can be explained by the fact that, when their actions are visible, people worry about others' approval [@Rege2004] and that, in non-anonymous situations, contributions can be a self-presentation strategy, because being perceived as pro-social can increase one's social status [@VanVugt2010a]. There are therefore more social incentives to cooperate in non-anonymous, rather than anonymous, situations.

Moreover, this effect is displayed especially in situations when individuals face a cost when cooperating and know that they will not get any personal benefit. For instance, @Butz2021 demonstrated that, in a public good game where a part of contributions was donated to a charity association, disclosing identities increased cooperation, but only if the donation was deducted from participants' payoff. In the same direction, @VanVugt2010a found that, when actions were public, contributions were higher if the public good was already attained or unattainable, and therefore if participants did not have a personal interest in cooperating.

Reducing anonymity in cooperative interactions can be very effective to encourage people to contribute more. A possible way to do this is to reduce group's size. In fact, according to @Romano2016, larger groups lead to de-individuation, because it is more difficult to understand who is defecting. Therefore, it is easier for members to feel anonymous and to perceive the social punishments associated with defection as less probable. Moreover, communication is easier in smaller groups [@VonBorgstede2018]. This assumption is confirmed experimentally by @Jiang2021, who showed than in smaller groups people contributed more often to a public good.

Another effective method to reduce anonymity is to make people publicly promise to cooperate before the cooperative interaction. For instance, in a laboratory experiment modeling a common pool resource, people who publicly committed exploited significantly less the resource than people who committed in private, even if after committing they were not restricted in their harvesting possibilities [@Mosler1993]. This results was also confirmed by @Przepiorka2020, who showed that non-binding promises increased cooperation. Moreover, even a compulsory commitment has a beneficial effect on cooperation, even if the most efficient groups are those where all participants signed voluntary the pledge [@Koessler2021a].

In conclusion, reducing anonymity during interactions is essential to improve cooperation, because this can facilitate the implementation of others cooperation-enhancing techniques [@Wang2017]. Moreover, there are several low-cost method to make situations non-anonymous.

#

>>>>

-   **Shorten OK, see how organize the table (putting all the details of the original paper?)**

@@

Five individual factors that can impact cooperation emerged from the litterature review: gender, social status, group identification, values and personality traits, and trust. In the following sections, the principal effects of these factors on cooperation are described. Moreover, an overview of the moderators of each effect is provided. More details about moderators are included in the Appendix A.

@@

Social identity has a strong, positive effect on cooperation: the more individuals identify with a group, the more they cooperate with other members of the group [@VanVugt2009], for instance by providing more voluntary activities for the collective good [@Noonan2016]. In fact, humans have a very developed sense of belonging to social groups, and when they strongly identify with a social group, they are more concerned about their group's well-being and reputation [@VanVugt2009] and they value more group-related goal [@Arora2016]. A strong identification also maintains high levels of cooperation over time [@Zhang2019]. Finally, further confirmations of the strength of this effect are the fact that it is the same for individualist and collectivist people [@Chen2007] and that group identification can be easily implemented in a laboratory experiment and still have a strong effect on cooperation [@Wit1992].

@@

Intragroup cooperation can be enhanced via a competition between different groups, because in such a context individuals' and groups' interests are aligned [@Puurtinen2009]. Moreover, a study showed that intergroup competition increased cooperation, while individual, dispositional competitiveness had the opposite effect [@Nockur2020]. This demonstrated that the increase of cooperation via competition is due to group membership factors, and not to competition in itself.

Despite the beneficial effects of inducing an intergroup competition, this could be harmful in real-life situations, like for instance in case of resources shared between several groups. In fact, there is a greater risk of over-exploitation if an intergroup competition is present [@VanVugt2009]. To maintain the positive effect of group identification while eliminating the risks associated to the competition, it is useful to prime a superordinate identity [@VanVugt2009]. This can be done by making salient the common characteristics of the different groups rather than their differences [@Kramer1984] or by inducing the feeling of a common fate [e.g. a collective risk, @Zhang2019].

@@

On the organizational level, four factors have emerged from the literature review: incentives, communication, social norms and anonymity. As for the individual factors, the main effects of each factor are detailed in the next sections, as well as an overview of possible moderators. More details can be found in Appendix A.

@@

Cooperation can be modified by introducing incentives, who can be either punishments or rewards[^1]. In the literature, punishment has received a greater attention than rewards [@Balliet2011a; @VanDijk2015], and its positive effect on cooperation has been widely demonstrated. In fact, when participants can punish other members for not contributing equally cooperation s higher than when this is not possible [e.g. @Balliet2011a; @Pfattheicher2018]. Moreover, the presence of a punishment system increases the perception of fairness [@Nockur2021].

@@

Several factors moderate this main effect. Punishments are more effective if their implementation is (moderately) costly rather than free [@Balliet2011a; @Egas2008], if the members are friends rather than strangers, if they know that they will interact repeatedly [@Balliet2011a], and if only cooperators have the power to punish defectors [@Xiao2016]. Also, a punishment system increase cooperation mostly in groups where all members have the same income [@Kingsley2016], because in heterogeneous groups a form of solidarity between low-status members arises and encourages free-riding [@Chen2022]. Finally, a plain situation is needed to have a stronger effect of punishment: the link between behaviour and outcome must be clear, without random factors affecting the final outcome [@Xiao2016], and the punishment must be highly effective [@Egas2008].

At the same time, punishment must be applied with carefulness. For instance, the most successful groups are those who punish with caution and not in early stages of the interaction, and who have a norm of when and how punishment should be applied [@Rockenbach2019]. Moreover, punishment has also two negative sides. First, it could entails a second-order social dilemma, since cooperative participants can free-ride on the effort of other cooperators who punish non-cooperators [@Milinski2012]. Secondly, when all participants can punish independently of their contributions, anti-social punishment (i.e. the punishment of cooperators by non-cooperators) can arise [@Xiao2016]. A solution could be to implement a system of democratic punishment, where a fine is given only if a majority of participants agree. In fact, with this kind of system the punishment level is lower, while final payoff and cooperation rates are higher [@Pfattheicher2018].

On the other side, rewarding systems are less investigated. Generally, they have a positive effect on cooperation [@Balliet2011a], but it is not clear whether they are more efficient than punishment [as found by @Chen2022], less efficient [see @Milinski2012] or equivalent [see @Balliet2011a]. Similarly to punishments, rewards have a stronger effect if their implementation is costly and during iterated, rather than single, interactions [@Balliet2011a].

To conclude, incentives can enhance cooperation, but they have some negative sides. First, they reduce participants' intrinsic motivation to cooperate [@VanDijk2015; @Schlosser2018] by framing the situation as an economic, rather than ethical, dilemma [@VanVugt2009]. Secondly, incentives undermine participants' trust in other members' internal motivation to cooperate. Therefore, if they are removed, cooperation drops [@Irwin2014; @Mulder2006].

Finally, incentives are often studied in economic formats, but they can also be social. In fact, expecting honour or shame (after cooperating or defecting, respectively), or expressing disapproval toward defectors can enhance cooperation [@Jacquet2011; @Nelissen2013]. Even an ostracism system where participants can vote to exclude other members is very effective [@Maier-Rigaud2010]. In conclusion, social sanctions can be even more efficient than economic ones, because cooperation do not drop after their removal [@Nelissen2013].

@@

Communication between members has a global positive effect on cooperation [e.g. @Balliet2010], because it allows to shape a normative content [@Janssen2014; @Bicchieri2002] and to reach a consensus about the best strategy [@Hopthrow2005]. In fact, when communication is used to foster a common social identity, promote cooperative social norms and ensure that every member publicly commits, communication is higher [@Adams2022].

Concerning the moderation effects, oral communication is more powerful than written messages, and communication has a stronger impact in large, rather than small, groups [@Balliet2010]. Moreover, if communicating requires a personal cost, its effect is weaker than if communication is free [@Janssen2014]. The content of the message is also essential: to be effective, communication must allow participants to become aware of the problem, identify different strategies and agree on the one to use. However, casual chit-chat also plays a very important role in increasing cooperation [@Koessler2021]. Finally, communication before or during the cooperative interaction have both a positive impact on cooperation [@Balliet2010], while knowing that it will be possible to communicate after the interaction does not enhance cooperation [@Kumakawa2013], and can even have a negative effect [@Torsvik2011].

A particular form of communication is gossip, i.e. the spread of reputational information about another person [@Feinberg2014]. Gossip increases cooperation by providing information about the past behaviour of potential partners, which helps to select the most cooperative members in a following interaction. Moreover, knowing that gossip is possible motivates people to cooperate more to avoid being the target of such gossip [@Feinberg2014]. Gossip can also have the pro-social function of protecting others from a potential exploitation [@Feinberg2012], and providing correct information while gossiping helps to build a positive reputation [@Giardini2021].

@@

Social norms can provide a guide on how to behave, especially when there is uncertainty about others' behaviour like in cooperative situations [@VonBorgstede2018]. Both injunctive (what others approve) and descriptive (what others do) norms are positively related to cooperative behaviours [@Thogersen2008]. Moreover, social norms help to maintain cooperation, because individuals who do not respect them are susceptible of social sanction [@Rege2004].

Social norms shape cooperation because, according to the appropriateness framework, when taking a decision people ask themselves: "What does a person like me do in a situation like this?" [@Arora2012]. If they define the situation as a cooperative task and strongly identify with the group, group's norms will be more accessible and they will have a stronger impact on behaviour [@VanLange2013]. Moreover, people have a general tendency to imitate what others do, even if others' actions do not have an impact on their personal gains [@Bardsley2005a].

Finally, social norms can be modified by participants' behaviour. For instance, the presence of consistent cooperators (i.e. people who always cooperate, independently from others' actions) creates a more cooperative social norm and thus increases cooperation [@Weber2008]. Moreover, participants can also shape social norms via communication [@Bicchieri2002].

@@

In general, in non-anonymous situations cooperation is significantly higher than in anonymous ones [@Wang2017]. In fact, in such situations there are more social incentives to cooperate, because people worry about others' approval [@Rege2004]. Moreover, several factors that can increase cooperation rely on a reduction of anonymity [@Wang2017]: for instance, punish defectors is impossible without knowing who they are. When behaviours are visible, cooperation can also be a self-presentation strategy to be perceived as pro-social and therefore to increase one's social status. In fact, the positive effect of non-anonymity is stronger in situations where cooperation bears a cost without offering a personal gain [@Butz2021; @VanVugt2010a].

Therefore, reducing anonymity in cooperative interaction could significantly improve cooperation. A possible way to do this is to reduce group's size, because larger groups lead to de-individuation and it is thus more difficult to identify defectors [@Romano2016]. In fact, cooperation levels are higher in small, rather than large, groups [@Jiang2021]. Anonymity can also be reduced by making people publicly promise to cooperate. Such a commitment increases significantly cooperation rates, even if after committing people are not forced to cooperate [@Mosler1993; @Przepiorka2020]. Moreover, even a compulsory commitment has a beneficial effect on cooperation [@Koessler2021a].
